% $Header: /Users/joseph/Documents/LaTeX/beamer/solutions/generic-talks/generic-ornate-15min-45min.en.tex,v 90e850259b8b 2007/01/28 20:48:30 tantau $

\documentclass[handout]{beamer}
\usepackage{pgfpages}
\pgfpagesuselayout{4 on 1}[a4paper,border shrink=5mm,landscape]


\mode<presentation>
{
  \usetheme{cambridge}
  %\usetheme{Warsaw}

  \setbeamertemplate{navigation symbols}{}
  \setbeamercovered{transparent}
  % or whatever (possibly just delete it)
}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{multicol}

%\usepackage{times}
%\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.


\title[HPC: An introduction] % (optional, use only with long paper titles)
{An Introduction to High Performance Computing}

%\subtitle
%{Presentation Subtitle} % (optional)

\author[SJ Rankin] % (optional, use only with lots of authors)
{Stuart Rankin\\ \texttt{sjr20@cam.ac.uk}}
%{F.~Author\inst{1} \and S.~Another\inst{2}}
% - Use the \inst{?} command only if the authors have different
%   affiliation.

\institute[Research Computing Services, University of Cambridge] % (optional, but mostly needed)
{Research Computing Services (http://www.hpc.cam.ac.uk/)\\
University Information Services (http://www.uis.cam.ac.uk/)}
% - Use the \inst command only if there are several affiliations.
% - Keep it simple, no one is interested in your street address.

\date[20/06/2019] % (optional)
{20th June 2019 / UIS Training}

\subject{Courses}


\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}<presentation>{Welcome}
\begin{itemize}
\item{Please sign in on the {\color{red}attendance sheet}.}
\item Please give your {\color{red}online feedback} at the end of the course:
      \url{http://feedback.training.cam.ac.uk/ucs/form.php}
\item{Keep your belongings with you.}
%\item Course files can be downloaded from:  \url{www.ucs.cam.ac.uk/training/files}
\item\alert{Please ask questions and let us know if you need assistance.}
\end{itemize}
\end{frame}

\section{Who are we?}
\begin{frame}{UIS: Research Computing Services}
Your trainers for today will be:\\
\begin{itemize}
\item{\alert{Stuart Rankin}\\{}\qquad\qquad Research Computing User Services}
\item{\alert{Mark Sharpley}\\{}\qquad\qquad Research Computing Platforms}
\item{We are generalists, but there is also the \alert{Research Software Engineering} team.}
\end{itemize}
\end{frame}

\section{Who are you?}
\begin{frame}{You may be $\ldots$}
  \begin{itemize}
  \item{Programmers (or not).}\pause
  \item{UNIX power users (or not).}\pause
  \item{Researchers wishing to run large, parallel code.}\pause
  \item{Researchers wishing to run many, non-parallel cases.}\pause
  \item{Researchers interested in big data, machine learning, AI.}\pause
  \item{Researchers requiring slightly more than an ordinary workstation.}\pause
  \item{\alert{Many different disciplines and requirements.}}
\end{itemize}
\end{frame}


\begin{frame}<presentation>{Plan of the Course}
\begin{description}
\item[Part 1:]{Basics}
\item[Part 2:]{Research Computing Services HPC}
\item[Part 3:]{Using HPC}
\end{description}
\end{frame}

\include{main}

\end{document}
